---
title: "Reseach Questions"
author: "Ulita Gilbert"
date: "3 May 2021"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library (janitor)
library(skimr)
library(plotly)
library(scales)
library(httr)
library(jsonlite)
library(lubridate)
```

# Introduction

In Part B of this project, I will further examine the effects of the pandemic on recent MCPS highschool graduates enrolled at Montgomery College. For the purposes of this study I will limit my dataset to MCPS students under the age of 20. These MCPS students will be divided further into subgroups based on Gender and Race. The datasets used in this part of my project have already been cleaned in Part A.   

In term year 2020, 74.5% of the full-time students graduated from MCPS high schools and were 20yrs old and younger.
52% of the part-time students were MCPS graduates who were 20yrs and younger.  
In term year 2021, 75% of the full-time students graduated from MCPS high schools and were 20yrs old and younger.
58% of the part-time students were MCPS graduates who were 20yrs and younger.  


# Data Dictionary

For the purposes of this Project the following variables and definitions are important.

**Terminology:**  
**Fall2019** refers to the incoming freshman cohort in Fall2019. This is term year 2020.    
**Fall2020** refers to the incoming freshman cohort in Fall2020. This is term year 2021.  

**Variables of Interest:**   
**term year:** incoming students in Fall 2019 are assigned to term year 2020. Incoming students in Fall 2020 are assigned to term year 2021.    
**hours_earned:** refers to credit hours the student has earned in their first Fall semester. This can include credits earned in Summer school second session (Summer 2) and AP credits earned in high school.    
**hours_attempted:** refers to credit  and non credit hours the student has attempted in their first Fall semester. This may include credits attempted in Summerschool second session - Summer 2.    
**full_part:** is the student full-time (FT) or part-time (PT). This classification is based on the students self reported information in the admissions application. Students are classified as full-time if they intend to take at least 12 credits.   
**major:** degree programme student is registered for or certificate&LR ( letter of recommendation.) All certificates and letters of recommendations have been grouped together.     
**hours_earned_rate:** Ratio of hours_earned/hours_attempted   
**age:** age of student at the start of program.    
**race:** racial classification of student. This is based on the IPEDS system. Foreign students are identified as foreign and not by their race/ethnicity.  
**sex:** gender classification of student.  
**high_school:** name of highschool student graduted from. Public High schools in Montgomery county are classified as MCPS.  
**pell:** Whether the student received a pell grant or not.  


# Data Wrangling
## Import Data 

```{r, include=FALSE}

#set directory
setwd('/Users/ulita/Documents/CAPSTONE205/PARTB')

df_Degrees<-read_csv("Degrees_Data.csv")
#df_MCPS20D<-read_csv("df_MCPS20D.csv")
#MCPS_list<-read_csv("MCPS_Data.csv")

```

Summary of Data and Types
```{r}

skim(df_Degrees)

```

Change Datatypes
```{r}
df_Degrees$u_number<- as.character(df_Degrees$u_number)
df_Degrees$term_year<- as.character(df_Degrees$term_year)
```


## Create DataFrame of students who graduated MCPS high schools who are 20yrs and under .

Use the dataframe df_Degrees which has been cleaned in the initial data analysis.
Filter all MCPS students who are 20yrs and younger in age. 
```{r}
df_MCPS20D<-df_Degrees %>%                    
         filter(HS_classify=="MCPS")%>%    # filter degrees dataset to obtain students who graduated MCPS highschools
         filter(age=='18 - 20' | age =="< 18") # filter students who are 20yrs old and younger. 

```


# Demographics of Students who graduated from MCPS highschools and are 20yrs and younger.

## Full time versus Part-time Degree Students

Frequency of Students Part time versus Full tim: 2020 vs 2021
```{r}

# Number of students part time abnd full time  2020 vs 2021
ggplot(data=df_MCPS20D, aes(x=full_part, fill=full_part)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=2,size=3)+
      facet_wrap(~term_year)+
      ggtitle("Number of Students Full time versus Part time")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   
```

There was a 5.98% decrease in full time students who graduated from MCPS highschools in term year 2021. There was a 6.74% decrease in part time students who graduated from MCPS.

Proportion of Students Full time versus Part time: 2020 vs 2021
```{r}

df_MCPS20D %>% 
    group_by(term_year) %>% 
    count(full_part) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = full_part, y = prop)) +
    geom_col(aes(fill = full_part), position = "dodge") +
    geom_text(aes(label = scales::percent(prop), 
                  y = prop, 
                  group = full_part),
              position = position_dodge(width = 0.9),
              vjust = 2,size=3)+
   facet_wrap(~term_year)+
      ggtitle("Proportion of Students Full time versus Part time")+
      ylab('Percentage')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   

```
In both term-year 2020 and term year 2021, approximately 2/3rds of the students are full time and 1/3 are part time. 

```{r}
# change in overall MCPS student population from 2020 to 2021

df_MCPS20D%>%
          group_by(term_year,full_part)%>%
          count(full_part)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(full_part)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```


Students who graduated from MCPS high schools, aged 20yrs and younger vs Full Freshman Cohort
```{r}
term_year<-c(2020,2020,2020,2020,2021,2021,2021,2021)
full_part<-c("FT","FT","PT","PT","FT","FT","PT","PT")
population<-c("Cohort","MCPS","Cohort","MCPS","Cohort","MCPS","Cohort","MCPS")
num<-c(2222,1655,1527,801,2088,1566,1286,747)

MC_data<-data_frame(term_year,full_part,population,num)

MC_data




ggplot(MC_data, aes(x=full_part,y=num,fill=population)) +
     geom_bar(position ="dodge",stat="identity")+
      geom_text(aes(label=num), position=position_dodge(width=0.9), vjust=-0.25)+

      facet_wrap(~term_year)+
     # scale_y_continuous(labels = scales::percent)+
      ggtitle("MCPS graduates 20yrs and younger vs Full Cohort")+
      ylab('')+
      xlab("")

 
   
```
In term year 2020, 74.5% of the full-time students graduated from MCPS high schools and were 20yrs old and younger.
52% of the part-time students were MCPS graduates who were 20yrs and younger.  
In term year 2021, 75% of the full-time students graduated from MCPS high schools and were 20yrs old and younger.
58% of the part-time students were MCPS graduates who were 20yrs and younger.  



## Race

Count of Race Groups
```{r}
ggplot(data=df_MCPS20D, aes(x=race, fill=race)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0,size=3)+
      facet_wrap(~term_year + full_part)+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())+
      ggtitle("Number of Students per a Race Group")+
      xlab("Race")+
      ylab("Frequency")


```


Percentage of each race group in Student Population
```{r}
df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    count(race) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = race, y = prop)) +
    geom_col(aes(fill = race), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.1), 
                  y = prop, 
                  group = race),
              position = position_dodge(width = 0.9),
              vjust = 0,size=3)+
    facet_wrap(~term_year + full_part)+
      ggtitle("Race Proportion of Students: Full time versus Part time")+
      ylab('Proportion ')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   

```



Full time student: Change in enrollment from 2020 to 2021 based on Race
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  race

df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,race)%>%
          count(race)%>%
          group_by(race)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Full time students: There was a 16.5% decline in asian students, 16.2% decline in African American students, a 9.1% decline in white students and 6.8% decline in foreign students. Hispanic students increased by 11.6%.

Part time student: Change in enrollment from 2020 to 2021 based on Race
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  race

df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,race)%>%
          count(race)%>%
          group_by(race)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Part time students: There was an 8.7% decrease in Asian students, a 26% decrease in foreign students, a 19.6% decrease in hispanic students. There was a 31.25% increase in white students and a 2.3% increase in african american students.

## Gender
Gender of Students 
```{r}

# Gender of students part time and full time  2020 vs 2021
ggplot(data=df_MCPS20D, aes(x=sex, fill=sex)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=1,size=3)+
      facet_wrap(~term_year+full_part)+
      ggtitle("Gender of Students: Full time versus Part time")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   
```

Calculate percentage change in full time student enrollment from 2020 to 2021 by  gender
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  gender

df_MCPS20D%>%
          filter(full_part=="FT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,sex)%>%
          count(sex)%>%
          group_by(sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```
Full time students: 14% decrease in attendance by male students. A 3.27% decrease in female students.


Calculate percentage change in part time student enrollment from 2020 to 2021 by  gender
```{r}
# calculate percentage change in part time student enrollment from 2020 to 2021 by  gender

df_MCPS20D%>%
          filter(full_part=="PT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,sex)%>%
          count(sex)%>%
          group_by(sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Part time: 9.5% decrease in female students. 1.5% decrease in male students. 



Gender and Race breakdown of full time students
```{r}
# Gender and Race of full time students  2020 vs 2021

df_MCPS20D%>%
      filter(sex %in% c("F","M"))%>%
      filter(full_part=="FT")%>%
      ggplot(., aes(x=race, fill=race)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, size=3)+
      facet_wrap(~term_year+sex)+
      ggtitle("Gender and Race of Full time Students")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())

#    theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())

```

Full time Student Enrollment Percentages trend by Gender and race

```{r}
# calculate percentage change in student enrollment from 2020 to 2021 by race and gender

# create data frames with counts of full time students by race and gender
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,race,sex)%>%
          count(sex)%>%
          group_by(race,sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```
Full time male students:  
30.82% decrease in asian males. 23.76% decrease in black/african american males. 22% screase in foreign males, 15.80% decrease in white males. 8.11% inscrease in Hispanic Males.  
Full time female students:    
33% decrease in foreign females. 14% increase  in Hispanic females. 

Part time Student Enrollment Percentages trend by Gender and race
```{r}
# calculate percentage change in student enrollment from 2020 to 2021 by race and gender

# create data frames with counts of full time students by race and gender
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,race,sex)%>%
          count(sex)%>%
          group_by(race,sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```
Part time male students:  
36.67% decrease in asian male students. 39.66% increase in white males. 14.28% decrease in hispanic males. 2% decrease in black/african american males. 30.95% decrease in foreign males. 
Part time female students:  
18.91% increase in asian female students. 6% increase in black/african american females. 19.16% decrease in hispanic females. 24% increase in white females. 23% decrease in foreign females. 



## Pell Grant


```{r}

# Pell Grant
ggplot(data=df_MCPS20D, aes(x=pell_grant, fill=pell_grant)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=3, size=3)+
      facet_wrap(~term_year+full_part)+
      ggtitle("Pell grant")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   
```

Proportion of Students receiving pell grants
```{r}
df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    count(pell_grant) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = pell_grant, y = prop)) +
    geom_col(aes(fill = pell_grant), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.1), 
                  y = prop, 
                  group = pell_grant),
              position = position_dodge(width = 0.9),
              vjust = 0,size=3)+
    facet_wrap(~term_year + full_part)+
      ggtitle("Proportion of Students receiving Pell Grants")+
      ylab('Proportion ')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   

```

## Majors

Overall Majors trend

Count of Majors in Full time students in 2020
```{r}
z1<- df_MCPS20D%>%
      filter(full_part=="FT" &term_year =="2020")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Full-time Students in 2020  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z1 + coord_flip()
```

Count of Majors in Full time students in 2021
```{r}
z13<- df_MCPS20D%>%
      filter(full_part=="FT" &term_year =="2021")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Full-time Students in 2021  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z13 + coord_flip()
```


calculate percentage change in full time student majors from 2020 to 2021 
```{r}
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,major)%>%
          count(major)%>%
          group_by(term_year)%>%
          group_by(major)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)
```
Fulltime: There was a 63% increase in health Sciences and a 28% decrease in general studies. 10.67% increase in business.


Count of Majors in Part time students in 2020
```{r}
z11<- df_MCPS20D%>%
      filter(full_part=="PT" &term_year =="2020")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Part-time Students in 2020  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z11 + coord_flip()
```
Count of Majors in Part time students in 2021
```{r}
z12<- df_MCPS20D%>%
      filter(full_part=="PT" &term_year =="2021")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Part-time Students in 2021  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z12 + coord_flip()
```

calculate percentage change in part time student majors from 2020 to 2021 
```{r}
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,major)%>%
          count(major)%>%
          group_by(term_year)%>%
          group_by(major)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)
```
Part-time: 21.4% decrease in general studies. A 5.6% increase in Health Sciences. 


## High Schools

### Full time Student

Breakdown of Highschools Full time students in term year 2020 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="FT" & term_year=="2020")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

Breakdown of Highschools Full time students in term year 2021 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="FT" & term_year=="2021")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by MCPS highschool
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)%>%
          arrange(desc(pct_change))
     
  
```

```{r}
v1<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="FT" & term_year=="2020")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools full time students graduated in term year 2020 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v1+ coord_flip()  
```

```{r}
v1<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="FT" & term_year=="2021")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools full time students graduated in term year 2021 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v1+ coord_flip()  
```


### Part time Student

Breakdown of Highschools Part time students in term year 2020 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="PT" & term_year=="2020")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

Breakdown of Highschools Part time students in term year 2021 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="PT" & term_year=="2021")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by MCPS highschool
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)%>%
          arrange(desc(pct_change))
     
  
```

```{r}
v3<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="PT" & term_year=="2020")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools Part time students graduated in term year 2020 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v3+ coord_flip()  
```

```{r}
v4<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="PT" & term_year=="2021")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools Part time students graduated in term year 2021 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v4 + coord_flip()  
```



# Hours Attempted


Boxplots of hours_attempted by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(hours_attempted))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```


 Boxplots of hours_attempted by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_attempted))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of hours_attempted by year by Part time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_attempted))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

There are not many outliers in the part time student groups. Term year 2021 seems to have more outliers on the upper end. 

Density plot of hours_attempted by year
```{r}
ggplot(df_MCPS20D, aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("Hours attempted") +
  ylab( "Density")+
   ggtitle(" Hours Attempted by Full-time Students vs Part-time Students")

```



 Hours attempted by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours attempted") +
  ylab( "Density") +
  ggtitle(" Hours Attempted by Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_attempted)[1],
            Q1 = fivenum(hours_attempted)[2],
            median = fivenum(hours_attempted)[3],
            Q3 = fivenum(hours_attempted)[4],
            max = fivenum(hours_attempted)[5],
            mean= mean(hours_attempted),
            sd = sd(hours_attempted))

```
Full time students:



 Hours attempted by part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours attempted") +
  ylab( "Density")+
   ggtitle(" Hours Attempted by Part-time Students")

```
 
 Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_attempted)[1],
            Q1 = fivenum(hours_attempted)[2],
            median = fivenum(hours_attempted)[3],
            Q3 = fivenum(hours_attempted)[4],
            max = fivenum(hours_attempted)[5],
            mean= mean(hours_attempted),
            sd = sd(hours_attempted))

```

 

# Hours Earned

Boxplots of Hours Earned by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(hours_earned))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```




 Boxplots of hours_earned by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of hours_earned by year by Part time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

There are not many outliers in the part time student groups. Term year 2021 seems to have more outliers on the upper end. 

Density plot of hours_earned by year
```{r}
ggplot(df_MCPS20D, aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("Hours Earned") +
  ylab( "Density")+
  ggtitle(" Hours Earned by Full-time vs Part-time Students")

```



Hours_earned by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" Hours Earned by Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_earned)[1],
            Q1 = fivenum(hours_earned)[2],
            median = fivenum(hours_earned)[3],
            Q3 = fivenum(hours_earned)[4],
            max = fivenum(hours_earned)[5],
            mean= mean(hours_earned),
            sd = sd(hours_earned))

```


hours_earned by part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" Hours Earned by Part-time Students")

```
 

Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_earned)[1],
            Q1 = fivenum(hours_earned)[2],
            median = fivenum(hours_earned)[3],
            Q3 = fivenum(hours_earned)[4],
            max = fivenum(hours_earned)[5],
            mean= mean(hours_earned),
            sd = sd(hours_earned))

```


# GPA
 
Boxplots of GPA by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(mc_gpa))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```




 Boxplots of GPA by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(mc_gpa))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of GPA by year by Part time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(mc_gpa))+
  ggtitle("GPA of Part-time students by Race")+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```



Density plot of GPA by year
```{r}
ggplot(df_MCPS20D, aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("GPA") +
  ylab( "Density")+
  ggtitle(" GPA by Full-time vs Part-time Students")

```



GPA by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" GPA of Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(mc_gpa)[1],
            Q1 = fivenum(mc_gpa)[2],
            median = fivenum(mc_gpa)[3],
            Q3 = fivenum(mc_gpa)[4],
            max = fivenum(mc_gpa)[5],
            mean= mean(mc_gpa),
            sd = sd(mc_gpa))

```


GPA of Part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" GPA of Part-time Students")

```
 

Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(mc_gpa)[1],
            Q1 = fivenum(mc_gpa)[2],
            median = fivenum(mc_gpa)[3],
            Q3 = fivenum(mc_gpa)[4],
            max = fivenum(mc_gpa)[5],
            mean= mean(mc_gpa),
            sd = sd(mc_gpa))

```
 
 
 # Hours Earned Rate
 
 Density plot of Hours Earned Rate by year
```{r}
ggplot(df_MCPS20D, aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.3) +
  facet_wrap(~full_part)+
  xlab("Hours Earned Rate") +
  ylab( "Density")+
  xlim(0,1)
```

Boxplots of Hours Earned Rate  of Full time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned_rate))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```



Boxplots of Hours Earned Rate  of Part time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned_rate))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```


GPA of full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" GPA of Full-time Students")

```


GPA of part time students 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" GPA of Part-time Students")

```


As a GPA of 2 is required to remain in good standing at the college, one needs to check the proportion of students with GPA >=2.


```{r}

df_MCPS20D%>% 
            group_by(race,full_part,term_year)%>%
            count(mc_gpa>=2)%>%
            arrange(term_year,.by_group=TRUE)%>%
            mutate(pct_pass= n/(n+lag(n))*100)


```



# Distribution of Variables and Correlation 

Distribution of Variables and Correlation : Full time Students 2020
```{r,message=FALSE}

library(GGally)
# plot distributions and correlation of variables
df_MCPS20D%>% filter(term_year=="2020")%>%
              filter(full_part=="FT")%>%
              ggpairs(., columns = c("hours_attempted","hours_earned", "mc_gpa","hours_earned_rate"))

```

Distribution of Variables and Correlation : Full time Students 2021
```{r,message=FALSE}

library(GGally)
# plot distributions and correlation of variables
df_MCPS20D%>% filter(term_year=="2021")%>%
              filter(full_part=="FT")%>%
              ggpairs(., columns = c("hours_attempted","hours_earned", "mc_gpa","hours_earned_rate"))

```


Distribution of Variables and Correlation : Part time Students 2020
```{r,message=FALSE}

library(GGally)
# plot distributions and correlation of variables
df_MCPS20D%>% filter(term_year=="2020")%>%
              filter(full_part=="PT")%>%
              ggpairs(., columns = c("hours_attempted","hours_earned", "mc_gpa","hours_earned_rate"))

```

Distribution of Variables and Correlation : Part time Students 2021
```{r,message=FALSE}

library(GGally)
# plot distributions and correlation of variables
df_MCPS20D%>% filter(term_year=="2021")%>%
              filter(full_part=="PT")%>%
              ggpairs(., columns = c("hours_attempted","hours_earned", "mc_gpa","hours_earned_rate"))

```
