---
title: "Reseach Questions"
author: "Ulita Gilbert"
date: "4/14/2021"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library (janitor)
library(skimr)
library(plotly)
library(scales)
library(httr)
library(jsonlite)
library(lubridate)
```

# Introduction

In this part of my project I will refine my research questions. I will further examine the effects of the pandemic on recent MCPS highschool graduates enrolled at Montgomery College. For the purposes of this study I will limit my dataset to MCPS students under the age of 20. These MCPS students will be divided further into subgroups based on Gender and Race. The datasets used in this part of my project have already been cleaned in my initial data analysis. Outliers have not been removed. I will conduct my statistical analysis with and without the outliers. 


# Data Dictionary

For the purposes of this Project the following variables and definitions are important.

The population in this dataset is the incoming cohort of students in Fall of 2019 and 2020. These students are first time degree or certificate seekers and have no prior tertiary education. They may have earned AP credits in highschool.

**Fall2019** refers to the incoming freshman cohort in Fall2019. This is term year 2020.  
**Fall2020** refers to the incoming freshman cohort in Fall2020. This is term year 2021.

**Variables of Interest:** 
**term year ** Incoming students in Fall2019 are assigned to term year 2020. Incoming students in Fall 2020 are assigned to term year 2021.  
**hours_earned:** refers to credit hours the student has earned in their first Fall semester ( this can include credits earned in Summer school second session- Summer 1 and AP credits earned in high school).  
**hours_attempted:** refers to credit  and non credit hours the student has attempted in their first Fall semester  ( this may include credits attempted in Summerschool second session - Summer 1).  
**full_part:** is the student full-time (FT) or part-time (PT). Part time students are registered in less than 12 credit hours. Full-time students take at least 12 credits.
**major:** degree programme student is registered for or certificate&LR ( letter of recommendation.) All certificates and letters of recommendations have been grouped together.   
**hours_earned_rate:** Ratio of hours_earned/hours_attempted 
**age:** Age of student at start of program.  
**race:** Racial classification of student.
**sex:** Gender classification of student.
**high_school:** Name of highschool student graduted from. Public High schools in Montgomery county are classified as MCPS.
**pell:** Whether the student receives a pell grant or not. 


# Data Wrangling
## Import Data 

```{r, include=FALSE}
#import data for Fall 2016 to Fall 2020
#set directory
setwd('/Users/ulita/Documents/CAPSTONE205/Capstone/Capstone_Code/Statistical_Analysis')

df_Degrees<-read_csv("Degrees_Data.csv")
#df_MCPS20D<-read_csv("df_MCPS20D.csv")
#MCPS_list<-read_csv("MCPS_Data.csv")

```

Summary of Data and Types
```{r}

skim(df_Degrees)

```

Change Datatypes
```{r}
df_Degrees$u_number<- as.character(df_Degrees$u_number)
df_Degrees$term_year<- as.character(df_Degrees$term_year)
```


## Create DataFrame of students who graduated MCPS high schools who are 20yrs and under .

Use the dataframe df_Degrees which has been cleaned in the initial data analysis.
Filter all MCPS students who are 20yrs and younger in age. 
```{r}
df_MCPS20D<-df_Degrees %>%                    
         filter(HS_classify=="MCPS")%>%    # filter degrees dataset to obtain students who graduated MCPS highschools
         filter(age=='18 - 20' | age =="< 18") # filter students who are 20yrs old and younger. 

```


# Demographics of Students who graduated from MCPS highschools and are 20yrs and younger.

## Full time versus Part-time Degree Students

Frequency of Students Part time versus Full tim: 2020 vs 2021
```{r}

# Number of students part time abnd full time  2020 vs 2021
ggplot(data=df_MCPS20D, aes(x=full_part, fill=full_part)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=2,size=3)+
      facet_wrap(~term_year)+
      ggtitle("Number of Students Full time versus Part time")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   
```

```{r}
# change in overall MCPS student population from 2020 to 2021

df_MCPS20D%>%
          group_by(term_year,full_part)%>%
          count(full_part)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(full_part)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

There was a 5.98% decrease in full time students who graduated from MCPS highschools in term year 2021. There was a -6.74% decrease in part time students who graduated from MCPS.

## Race

Count of Race Groups
```{r}
ggplot(data=df_MCPS20D, aes(x=race, fill=race)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0,size=3)+
      facet_wrap(~term_year + full_part)+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())+
      ggtitle("Number of Students per a Race Group")+
      xlab("Race")+
      ylab("Frequency")


```


Full time student: Change in enrollment from 2020 to 2021 based on Race
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  race

df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,race)%>%
          count(race)%>%
          group_by(race)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Full time students: There was a 16.5% decline in asian students, 16.1% decline in African American students, a 9.1% decline in white students and 6.8% decline in foreign students. 
Hispanic students increased by 11.6%.

Part time student: Change in enrollment from 2020 to 2021 based on Race
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  race

df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,race)%>%
          count(race)%>%
          group_by(race)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Part time students: There was an 8.7% decrease in Asian students, a 26% decrease in foreign students, 2.3% increase in african american students and a 19.6% decrease in hispanic students. There was a 31.25% increase in white students. 

## Gender
Gender of Students 
```{r}

# Gender of students part time and full time  2020 vs 2021
ggplot(data=df_MCPS20D, aes(x=sex, fill=sex)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=1,size=3)+
      facet_wrap(~term_year+full_part)+
      ggtitle("Gender of Students: Full time versus Part time")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())
   
```

Calculate percentage change in full time student enrollment from 2020 to 2021 by  gender
```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by  gender

df_MCPS20D%>%
          filter(full_part=="FT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,sex)%>%
          count(sex)%>%
          group_by(sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```
Full time students: 14% decrease in attendance by male students. A 3.27% decrease in female students.


Calculate percentage change in part time student enrollment from 2020 to 2021 by  gender
```{r}
# calculate percentage change in part time student enrollment from 2020 to 2021 by  gender

df_MCPS20D%>%
          filter(full_part=="PT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,sex)%>%
          count(sex)%>%
          group_by(sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```

Part time: 9.5% decrease in female students. 1.5% decrease in male students. 



Gender and Race breakdown of full time students
```{r}
# Gender and Race of full time students  2020 vs 2021

df_MCPS20D%>%
      filter(sex %in% c("F","M"))%>%
      filter(full_part=="FT")%>%
      ggplot(., aes(x=race, fill=race)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, size=3)+
      facet_wrap(~term_year+sex)+
      ggtitle("Gender and Race of Full time Students")+
      ylab('Frequency')+
      xlab("")+
      theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())

#    theme(axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank())

```

Full time Student Enrollment Percentages trend by Gender and race

```{r}
# calculate percentage change in student enrollment from 2020 to 2021 by race and gender

# create data frames with counts of full time students by race and gender
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,race,sex)%>%
          count(sex)%>%
          group_by(race,sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```


Part time Student Enrollment Percentages trend by Gender and race
```{r}
# calculate percentage change in student enrollment from 2020 to 2021 by race and gender

# create data frames with counts of full time students by race and gender
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          filter(sex=="F"|sex =="M")%>%
          group_by(term_year,race,sex)%>%
          count(sex)%>%
          group_by(race,sex)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100) 
```


## Majors

Overall Majors trend

Count of Majors in Full time students in 2020
```{r}
z1<- df_MCPS20D%>%
      filter(full_part=="FT" &term_year =="2020")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Full-time Students in 2020  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z1 + coord_flip()
```

Count of Majors in Full time students in 2021
```{r}
z13<- df_MCPS20D%>%
      filter(full_part=="FT" &term_year =="2021")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Full-time Students in 2021  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z13 + coord_flip()
```


calculate percentage change in full time student majors from 2020 to 2021 
```{r}
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,major)%>%
          count(major)%>%
          group_by(term_year)%>%
          group_by(major)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)
```

Count of Majors in Part time students in 2020
```{r}
z11<- df_MCPS20D%>%
      filter(full_part=="PT" &term_year =="2020")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Part-time Students in 2020  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z11 + coord_flip()
```
Count of Majors in Part time students in 2021
```{r}
z12<- df_MCPS20D%>%
      filter(full_part=="PT" &term_year =="2021")%>%
       ggplot(., aes(x=major, fill=major)) +
      geom_bar() +
      geom_text(stat='count', aes(label=..count..), vjust=0, hjust=0, size =3)+
      ggtitle("Majors of Part-time Students in 2021  ")+
      xlab("Major")+
      ylab("Frequency")+
    theme(legend.position = "none") 
       
z12 + coord_flip()
```

calculate percentage change in part time student majors from 2020 to 2021 
```{r}
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,major)%>%
          count(major)%>%
          group_by(term_year)%>%
          group_by(major)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)
```

## High Schools

### Full time Student

Breakdown of Highschools Full time students in term year 2020 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="FT" & term_year=="2020")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

Breakdown of Highschools Full time students in term year 2021 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="FT" & term_year=="2021")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by MCPS highschool
df_MCPS20D%>%
          filter(full_part=="FT")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)%>%
          arrange(desc(pct_change))
     
  
```

```{r}
v1<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="FT" & term_year=="2020")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools full time students graduated in term year 2020 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v1+ coord_flip()  
```

```{r}
v1<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="FT" & term_year=="2021")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools full time students graduated in term year 2021 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v1+ coord_flip()  
```


### Part time Student

Breakdown of Highschools Part time students in term year 2020 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="PT" & term_year=="2020")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

Breakdown of Highschools Part time students in term year 2021 attended in MCPS
```{r}
df_MCPS20D%>%
          filter(full_part=="PT" & term_year=="2021")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          mutate(total_pop =sum(n))%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_pop= (n/total_pop*100))%>%
          arrange(desc(pct_pop))
  
```

```{r}
# calculate percentage change in full time student enrollment from 2020 to 2021 by MCPS highschool
df_MCPS20D%>%
          filter(full_part=="PT")%>%
          group_by(term_year,high_school)%>%
          count(high_school)%>%
          group_by(term_year)%>%
          group_by(high_school)%>%
          arrange(term_year,.by_group=TRUE)%>%
          mutate(pct_change= (n-lag(n))/lag(n)*100)%>%
          arrange(desc(pct_change))
     
  
```

```{r}
v3<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="PT" & term_year=="2020")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools Part time students graduated in term year 2020 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v3+ coord_flip()  
```

```{r}
v4<- df_MCPS20D %>% 
    group_by(term_year,full_part) %>% 
    filter(full_part=="PT" & term_year=="2021")%>%
    count(high_school) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = high_school, y = prop)) +
    geom_col(aes(fill=high_school), position = "dodge") +
    geom_text(aes(label = scales::percent(prop,0.5), 
                  y = prop, 
                  group = high_school),
              position = position_dodge(width = 0.9),
              vjust = 0, size=3, hjust=0)+
  #  facet_wrap(~term_year )+
      ggtitle("High schools Part time students graduated in term year 2021 graduated")+
      ylab('Proportion ')+
      xlab("")+
      theme(legend.position = "none", axis.text.x=element_blank(),strip.background = element_blank(),panel.grid = element_blank()) 
  
v4 + coord_flip()  
```




# Statistical Analysis with Outliers

For the purposes of this analysis I will run the analysis first with outliers and then after removing outliers.


## Hours Attempted


Boxplots of hours_attempted by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(hours_attempted))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```

Students who register for more than 18 credits require special permission from the department.
Further more a full time student is classified as someone who is enrolled in 12 or more credits. A part time student is classified
as someone who is enrolled in less than 12 credits. However based on thge dataset, a number of full time students attempt less than 12 credits and large
a number of part time students attempt more than 12 hours.



 Boxplots of hours_attempted by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_attempted))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of hours_attempted by year by Part time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_attempted))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

There are not many outliers in the part time student groups. Term year 2021 seems to have more outliers on the upper end. 

Density plot of hours_attempted by year
```{r}
ggplot(df_MCPS20D, aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("Hours attempted") +
  ylab( "Density")+
   ggtitle(" Hours Attempted by Full-time Students vs Part-time Students")

```



 Hours attempted by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours attempted") +
  ylab( "Density") +
  ggtitle(" Hours Attempted by Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_attempted)[1],
            Q1 = fivenum(hours_attempted)[2],
            median = fivenum(hours_attempted)[3],
            Q3 = fivenum(hours_attempted)[4],
            max = fivenum(hours_attempted)[5],
            mean= mean(hours_attempted),
            sd = sd(hours_attempted))

```


 Hours attempted by part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_attempted, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours attempted") +
  ylab( "Density")+
   ggtitle(" Hours Attempted by Part-time Students")

```
 
 Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_attempted)[1],
            Q1 = fivenum(hours_attempted)[2],
            median = fivenum(hours_attempted)[3],
            Q3 = fivenum(hours_attempted)[4],
            max = fivenum(hours_attempted)[5],
            mean= mean(hours_attempted),
            sd = sd(hours_attempted))

```

 

## Hours Earned

Boxplots of Hours Earned by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(hours_earned))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```




 Boxplots of hours_earned by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of hours_earned by year by Part time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

There are not many outliers in the part time student groups. Term year 2021 seems to have more outliers on the upper end. 

Density plot of hours_earned by year
```{r}
ggplot(df_MCPS20D, aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("Hours Earned") +
  ylab( "Density")+
  ggtitle(" Hours Earned by Full-time vs Part-time Students")

```



Hours_earned by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" Hours Earned by Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_earned)[1],
            Q1 = fivenum(hours_earned)[2],
            median = fivenum(hours_earned)[3],
            Q3 = fivenum(hours_earned)[4],
            max = fivenum(hours_earned)[5],
            mean= mean(hours_earned),
            sd = sd(hours_earned))

```


hours_earned by part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" Hours Earned by Part-time Students")

```
 

Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(hours_earned)[1],
            Q1 = fivenum(hours_earned)[2],
            median = fivenum(hours_earned)[3],
            Q3 = fivenum(hours_earned)[4],
            max = fivenum(hours_earned)[5],
            mean= mean(hours_earned),
            sd = sd(hours_earned))

```


## GPA
 
Boxplots of GPA by year by MCPS students 20yrs and younger
```{r}
p11 = ggplot(df_MCPS20D, aes(mc_gpa))
p11 + geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~full_part)
```




 Boxplots of GPA by year by Full time  MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(mc_gpa))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```

Boxplots of GPA by year by Part time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(mc_gpa))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```



Density plot of GPA by year
```{r}
ggplot(df_MCPS20D, aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~full_part)+
  xlab("GPA") +
  ylab( "Density")+
  ggtitle(" GPA by Full-time vs Part-time Students")

```



GPA by full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" GPA of Full-time Students")

```

Fivenum Summary of Full time students
```{r}
df_MCPS20D%>% filter(full_part=="FT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(mc_gpa)[1],
            Q1 = fivenum(mc_gpa)[2],
            median = fivenum(mc_gpa)[3],
            Q3 = fivenum(mc_gpa)[4],
            max = fivenum(mc_gpa)[5],
            mean= mean(mc_gpa),
            sd = sd(mc_gpa))

```


GPA of Part time students 
 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(mc_gpa, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("Hours Earned") +
  ylab( "Density")+
   ggtitle(" GPA of Part-time Students")

```
 

Fivenum Summary of Part time students
```{r}
df_MCPS20D%>% filter(full_part=="PT")%>%
  group_by(race,term_year)%>%
  summarise(n = n(),
            min = fivenum(mc_gpa)[1],
            Q1 = fivenum(mc_gpa)[2],
            median = fivenum(mc_gpa)[3],
            Q3 = fivenum(mc_gpa)[4],
            max = fivenum(mc_gpa)[5],
            mean= mean(mc_gpa),
            sd = sd(mc_gpa))

```
 
 
 ## Hours Earned Rate
 
 Density plot of Hours Earned Rate by year
```{r}
ggplot(df_MCPS20D, aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.3) +
  facet_wrap(~full_part)+
  xlab("Hours Earned Rate") +
  ylab( "Density")+
  xlim(0,1)
```

Boxplots of Hours Earned Rate  of Full time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned_rate))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```



Boxplots of Hours Earned Rate  of Part time MCPS students 20yrs and younger
```{r}


df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
   ggplot(., aes(hours_earned_rate))+
   geom_boxplot(aes(colour = term_year)) +
       facet_wrap(~race)
```


Hours Earned Rate of full time students 
```{r}

df_MCPS20D%>%filter(full_part=="FT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" Hours Earned Rate of Full-time Students")

```


Hours Earned Rate of part time students 
```{r}

df_MCPS20D%>%filter(full_part=="PT")%>%
  filter(race=="White" |race=="Asian" |race=="Hispanic" |race=="Black / African Am." )%>%
  ggplot(., aes(hours_earned_rate, fill = term_year)) + geom_density(alpha = 0.2) +
  facet_wrap(~race)+
  xlab("GPA") +
  ylab( "Density")+
   ggtitle(" Hours Earned Rate of Part-time Students")

```


